Coherent optical association of a single molecule

### 2020-10-02 paper.tex
Update on Overleaf.

\begin{abstract}
  We report on coherent association of a single weakly-bound NaCs molecule in an optical tweezer
  through an optical Raman transition without the use of a Feshbach resonance. Our scheme borrows transition dipole moment while reducing photon scattering by selecting a deeply bound electronic excited intermediate state.
  Starting from two atoms in their relative motional ground state,
  we achieve optical transfer efficiency of $50 \%$ \todo{number}.
  The molecule has a (change to 8.8G) zero-field binding energy of $770 \mathrm{MHz}$ \todo{number}
  and lifetime up to $1 \mathrm{ms}$ \todo{number}. (center of mass motional state population.)
  % We demonstrate that coherent creation of a ground-state single molecule is possible, even without Feshbach resonance or narrow optical transition.
  This technique is general without relying on narrow excited state lines or Feshbach resonances and could allow a wider range of molecular species to be assembled atom-by-atom.
\end{abstract}

@@

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}A), where the initial atomic state, the target weakly-bound molecular state are coupled to an excited intermediate state by two photons, $\Omega_a$ and $\Omega_m$, with single photon detuning $ \Delta $, and two photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $c^3\Sigma^+$)
The transfer Raman Rabi Rate is $ \Omega_a\Omega_m / 2\Delta$~\cite{Wineland2003}. %, depends on both of these matrix elements and the single photon detuning~\cite{Wineland2003}.
Unlike for Raman transitions in atoms, the two matrix elements here, $ \Omega_m$ and $ \Omega_a$, are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the excited molecular state. The scattering rate therefore predominantly comes from the molecular state, $ \Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth. Because the energy difference of the atomic and the weakly bound states are small ($<$1~GHz) comparing to $\Delta$, the power of the Raman beams are chosen to be equal to gives the largest Raman Rabi rate at a fixed total power.) There is an additional factor of 2 in the scattering rate, since the molecular state scatters off of both Raman beams at roughly the same single photon detuning. Thus, the ratio between the Raman Rabi frequency and the scattering rate, $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth. To ensure a coherent process, larger detuning is preferred. However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency. With the multiple excited vibrational states present in a molecular potential, the total scattering rates and Rabi rates become a sum of the scattering rates and Rabi rates over all the states. There are also many possible intermediate states to choose from.

Pioneering experiments used weakly bound molecular excited states in the Raman transition to ensure a large Raman Rabi frequency. However, such choice of state also increases the scattering during the transfer process to render it incoherent, resulting in molecule loss~\cite{Wynar2000,Rom2004}. For a weakly bound molecular state, the spacing between states is generally smaller compared to deeply bound states and thus limits the maximum usable detuning and therefore the effectiveness of the scattering reduction by increasing the detuning. Furthermore, due to similarity in wavefunction size, the weakly bound molecular ground state and initial atomic state have strong coupling with the atomic excited state which contributes significantly to the total scattering rate. The rate of such a scattering process is proportional to $1/\delta_{thresh}^2$, where $\delta_{thresh}$ is the detuning from the dissociation threshold, and thus can also be made smaller by detuning away from the dissociation threshold.
%(Should we also mention scattering of the atomic initial state too? This should affect the Raman process as well)

To obtain a complete picture, we perform a full calculation of the Raman Rabi frequency and scattering rate at different detunings (full result in SM). As shown in Fig.~\ref{f-theory}B, the ratio of the Raman Rabi rate to scattering rate can be made better for more deeply bound states compared to weakly bound states.  %Since for a typical molecular potential the spacing between states
% decreases with increasing vibrational levels,
% a weakly bound excited molecular state
% This further increases the scattering rate for a weakly bond excited state.
% Based on these considerations, % a weakly bound excited molecular state may not be the best choice for the intermediate state in the Raman transfer. Thus,
% we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in Fig.~\ref{f-theory}B,
% we found a distinct optimal regime from what has been explored experimentally previously.
Namely, despite a smaller Raman Rabi frequency at deeply bound intermediate states, the Rabi frequency to scattering rate ratio is much higher as compared to a weakly bound intermediate state. As a result, we choose a deeply bound molecular excited state ($v'=0$ of $c^3\Sigma^+$) to drive the Raman transition. (mention the trend, also v=12 will work as similarly as v=0)

@@

% In addition to the intermediate state, the choice of the initial and final states are critical for both fundamental and technical reasons.  Thus, a small $ \Omega_m/\Omega_a $ ratio is desired to be able to drive this transition coherently at smaller detunings where the Raman Rabi frequency will be larger. In addition to this fundamental reason, there is also a technical advantage for having a $ \Omega_m/\Omega_a $ ratio to be as small( KW: I think small, but open question whether we should always talk about $\Omega_m/\Omega_a$ or $ \Omega_a/\Omega_m$, KN: maybe $ \Omega_a/\Omega_m$ is better because it sounds more natural that we want to make it large? but either way.) as possible. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser needs to be stabilized to better than the inverse of this ratio, $ \Omega_a/\Omega_m $, to fluctuate by less than a linewidth. This becomes technically easier when $ \Omega_m/\Omega_a $ is smaller.
% (for our choice of states, this mean intensity fluctuation of less than 1\% or 0.1\%. KW: I mentioned this in next paragraph, does it still need to be mentioned here?)

In addition to the intermediate state, the choice of the initial and final states is critical. As previously discussed, the ratio of the Raman Rabi rate to the scattering rate is proportional to the $ \Omega_a/\Omega_m $ ratio. The smaller this ratio is, the farther the laser needs to be detuned to achieve the same Raman Rabi rate to scattering rate ratio, which lowers the Raman Rabi rate. This ratio can be made larger by increasing the coupling of the atomic state with the excited molecular state. Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In addition to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (Fig.~\ref{f-theory}C). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, among the stable spin combinations, 4422 and 3311 both have small scattering lengths of $ a_{4422}= 30.4a_0, a_{3311} = 13.7a_0 $ respectively, but the 3332 combination has a large and negative scattering length of $ a_{3322} = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}. In addition to the increased atomic coupling, $ \Omega_a $, with the 3322 hyperfine combination, coupled channel calculations show that the most weakly bound molecular state that is predominantly in the 3322 spin combinationn also has reduced coupling, $ \Omega_m $, with the excited state when compared with the 4422 and 3311 bound states. Therefore, using an initial 3322 hyperfine combination results in a $ \Omega_a/\Omega_m$ ratio of about 0.05 instead of a ratio of about 0.003 with the 4422 or 3311 combinations. In addition to reducing how far the laser needs to be detuned from the transition, a larger $ \Omega_a/\Omega_m$ ratio also relaxes the intensity stability requirement, a key potential technical limitation. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser intensity needs to be stabilized to better than the inverse of this ratio to fluctuate by less than a linewidth. Thus, instead of a required intensity stability of $0.3 \% $ for the 4422 or 3311 combination, the 3322 combination only requires the intensity to be stabilized to $ 5\% $.  Thus, we choose the 3322 spin combination as our initial state and drive to the first bound state for the 3322 spin combination. (We might want to mention that those ratios are for 80 kHz spherical trap if we want to give more details about the coupled channel calculation anywhere...)

% In additional to the final and the excited state, it is also important to select the an initial atomic state in order to improve the coupling.
\begin{figure*}
  \includegraphics[height=4.5cm]{fig2.pdf}
  \caption{
    (A) Geometry and polarization of trap and Raman beam relative to the bias magnetic field.
    The tweezer and Raman beam is focused through an objective to define the location of the
    atoms and molecule.
    We use a bias B field of $B_0=8.8 G$ alone the tweezer polarization
    to define the quantization axis.
    As a result, the atoms experiences predominately $\pi$ polarization from the tweezer.
    (B) Molecule formation pulse sequence. The tweezer initially consists of only up leg power.
    When driving the Raman transition, the up leg power is smoothly ramped down and
    the down leg power ramped up over $10\mu s$ while maintaining the total power of the tweezer.
    This minimizes the heating on the atoms due to power fluctuation while maximizes the time
    with maximum Raman Rabi frequency when the up and down leg powers are equal.
    (C) Raman resonance from atomic state $Na(2,2)\ Cs(3,3)$ to the first molecular bound state
    using a $0.1\mathrm{ms}$ pulse.
    The full width half maximum (FWHM) of $8.43(81) \mathrm{kHz}$ of the resonance
    is consistent with the FWHM for a coherent $0.1\mathrm{ms}$ $8.0 \mathrm{kHz}$ $\pi$ pulse.
    (D) Raman pulse time scan on resonance. A decaying Rabi oscillation can be observed
    proving the coherence of the Raman transfer process.
    \label{f-raman}}
\end{figure*}

>>>>

\begin{abstract}
  We report on coherent association of a single weakly-bound NaCs molecule in an optical tweezer
  through an optical Raman transition without the use of a Feshbach resonance. Our scheme borrows transition dipole moment while reducing photon scattering by selecting a deeply bound electronic excited intermediate state.
  Starting from two atoms in their relative motional ground state,
  we achieve optical transfer efficiency of $50 \%$ \todo{number}.
  The molecule has a (change to 8.8G) zero-field binding energy of $770 \mathrm{MHz}$ \todo{number}
  and lifetime up to $1 \mathrm{ms}$ \todo{number}. (center of mass motional state population.)
  % We demonstrate that coherent creation of a ground-state single molecule is possible, even without Feshbach resonance or narrow optical transition.
  This technique is general without relying on narrow excited state lines or Feshbach resonances and could allow a wider range of molecular species to be assembled atom-by-atom. \textcolor{blue}{Yichao, can you update numbers in the abstract? also, can you update affiliations in the author list, Jeremy is Durham and Jon is Purdue}
\end{abstract}

@@

The essence of an optical Raman transfer can be illustrated using a three-level system (Fig.~\ref{f-theory}A), where the initial atomic state, the target weakly-bound molecular state are coupled to an excited intermediate state by two photons, $\Omega_a$ and $\Omega_m$, with single photon detuning $ \Delta $, and two photon detuning, $ \delta$.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $c^3\Sigma^+$)
The transfer Raman Rabi Rate is $ \Omega_a\Omega_m / 2\Delta$ accompanied by a photon scattering rate of $\Gamma_e \Omega / 4\Delta^2$...(sum of i terms\textcolor{red}{KW: I was intending to only mention the sum over terms at the end when we exit the 3 level system.})\textcolor{blue}{check factor of 2. KW: The 2 in the denominator is including cross coupling. 4 without cross coupling. Since cross coupling is mentioned later, maybe leave this with 4?}, where $ \Gamma_e $ is the excited state linewidth~\cite{Wineland2003}. %, depends on both of these matrix elements and the single photon detuning~\cite{Wineland2003}.
Unlike for Raman transitions in atoms, the two matrix elements here, $ \Omega_m$ and $ \Omega_a$, are greatly imbalanced %. Since $ \Omega_m $ is generally much larger than $ \Omega_a $
due to the small wavefunction overlap between the atomic state and the excited molecular state. The scattering rate therefore predominantly comes from the molecular state, $ \Gamma_e \Omega_m^2 / 2\Delta^2$, where $ \Gamma_e $ is the excited state linewidth. \textcolor{blue}{(blue text can potentially be moved to footnote) Furthermore, because the energy difference of the atomic and the weakly bound states are small ($<$1~GHz) comparing to $\Delta$, the power of the Raman beams are chosen to be equal to gives the largest Raman Rabi rate at a fixed total power. %There is 
An additional factor of 2 is included in the scattering rate because the molecular state scatters off of both Raman beams at roughly the same single photon detuning. } Thus, the ratio between the Raman Rabi frequency and the scattering rate, $ \Omega_a/\Omega_m \times \Delta/\Gamma_e $, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth. To ensure a coherent process, larger detuning is preferred while achieving a realistic Raman Rabi frequency. %However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency. 
With the multiple excited vibrational states present in a molecular potential, the total scattering rates and Rabi rates become a sum of the scattering rates and Rabi rates over all the states. %There are also many possible intermediate states to choose from.

Pioneering experiments used weakly bound molecular excited states in the Raman transition to ensure a large Raman Rabi frequency~\cite{Wynar2000,Rom2004}. However, such choice of state suffers from strong scattering of the nearby atomic excited state to render it incoherent, resulting in molecule loss. The scattering  is  proportional to $1/\delta_{thresh}^2$, where $\delta_{thresh}$ is the detuning from the dissociation threshold, and thus can  be made smaller by detuning away from the dissociation threshold.

 %For a weakly bound molecular state, the spacing between states is generally smaller compared to deeply bound states and thus limits the maximum usable detuning and therefore the effectiveness of the scattering reduction by increasing the detuning. 
%Furthermore, due to similarity in wavefunction size, the weakly bound molecular ground state and initial atomic state have strong coupling with the atomic excited state which contributes significantly to the total scattering rate. The rate of such a scattering process is proportional to $1/\delta_{thresh}^2$, where $\delta_{thresh}$ is the detuning from the dissociation threshold, and thus can also be made smaller by detuning away from the dissociation threshold.
%(Should we also mention scattering of the atomic initial state too? This should affect the Raman process as well)

To find the optimal excited intermediate state, we perform a  calculation of the Raman Rabi frequency and scattering rate at different detunings from the atomic threshold taking into account of all states of the $c^3\Sigma$ excited molecular state potential (full result in SM)~\cite{Liu2017} \textcolor{red}{KW: I have only done the calculations with c3Sigma only, cause the b3Pi is highly spin orbit coupled, and not necessarily trustworthy where we are. I can also include a citation to the paper with that potential?}. As shown in Fig.~\ref{f-theory}B, the ratio of the Raman Rabi rate to scattering rate can be made larger for more deeply bound states compared to weakly bound states at a cost of a smaller Raman Rabi frequency.  %Since for a typical molecular potential the spacing between states
% decreases with increasing vibrational levels,
% a weakly bound excited molecular state
% This further increases the scattering rate for a weakly bond excited state.
% Based on these considerations, % a weakly bound excited molecular state may not be the best choice for the intermediate state in the Raman transfer. Thus,
% we calculated both the Raman Rabi frequency and the scattering rate at different detunings (full result is shown in SM). As seen in Fig.~\ref{f-theory}B,
% we found a distinct optimal regime from what has been explored experimentally previously.
%Namely, despite a smaller Raman Rabi frequency at deeply bound intermediate states, the Rabi frequency to scattering rate ratio is much higher as compared to a weakly bound intermediate state. 
As a result, we choose a deeply bound molecular excited state ($v'=0$ of $c^3\Sigma^+$) as an intermediate state to drive the Raman transition. %(mention the trend, also v=12 will work as similarly as v=0)

@@

% In addition to the intermediate state, the choice of the initial and final states are critical for both fundamental and technical reasons.  Thus, a small $ \Omega_m/\Omega_a $ ratio is desired to be able to drive this transition coherently at smaller detunings where the Raman Rabi frequency will be larger. In addition to this fundamental reason, there is also a technical advantage for having a $ \Omega_m/\Omega_a $ ratio to be as small( KW: I think small, but open question whether we should always talk about $\Omega_m/\Omega_a$ or $ \Omega_a/\Omega_m$, KN: maybe $ \Omega_a/\Omega_m$ is better because it sounds more natural that we want to make it large? but either way.) as possible. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser needs to be stabilized to better than the inverse of this ratio, $ \Omega_a/\Omega_m $, to fluctuate by less than a linewidth. This becomes technically easier wh So en $ \Omega_m/\Omega_a $ is smaller.
% (for our choice of states, this mean intensity fluctuation of less than 1\% or 0.1\%. KW: I mentioned this in next paragraph, does it still need to be mentioned here?)

\textcolor{blue}{Main point here. Small $\Omega_a/\Omega_m$ ratio can be compensated by larger detuning, but Stark shift is problematic.} \textcolor{red}{KW: I agree Stark shift limitation is definitely pretty problematic, but I would also argue that larger detuning may not be able to save the day, because of the many vibrational states in a molecular potential, the lower Raman rabi frequency which results from detuning farther and possibly running into other limitations if you have to do that, such as the background off resonant scattering rate from near threshold states or technical background scattering.} \textcolor{blue}{yes, I agree. I want to layout the structure .... motivate for the better ratio first, than talk about choosing the state.} \textcolor{red}{KW: Okay, got it. Makes sense. }
In addition to the intermediate state, the choice of the initial and final states is also critical. As previously discussed, the ratio of the Raman Rabi rate to the scattering rate is proportional to the $ \Omega_a/\Omega_m $ ratio. The smaller this ratio is, the farther the laser needs to be detuned to achieve the same Raman Rabi rate to scattering rate ratio, which lowers the Raman Rabi rate. This ratio can be made larger by increasing the coupling of the atomic state with the excited molecular state. Due to the small size of the molecular wavefunction, the coupling between the ground atomic state and the excited molecular state is approximately proportional to the value of the relative atomic wavefunction at short distance within the molecular potential. In addition to the confinement, this value is related to the interaction between the two atoms. For states with a large scattering length (positive or negative), the phase shift in the relative wavefunction between the atoms can significantly increase the short range wavefunction (Fig.~\ref{f-theory}C). The increase in the coupling is proportional to (quote/cite Olive's equation?). For our system, among the stable spin combinations, 4422 and 3311 both have small scattering lengths of $ a_{4422}= 30.4a_0, a_{3311} = 13.7a_0 $ respectively, but the 3332 combination has a large and negative scattering length of $ a_{3322} = -693.8a_0 $ (interaction shift $\approx$ binding?)~\cite{Hood2019}. In addition to the increased atomic coupling, $ \Omega_a $, with the 3322 hyperfine combination, coupled channel calculations show that the most weakly bound molecular state that is predominantly in the 3322 spin combination also has reduced coupling, $ \Omega_m $, with the excited state when compared with the 4422 and 3311 bound states. Therefore, using an initial 3322 hyperfine combination results in a $ \Omega_a/\Omega_m$ ratio of about 0.05 instead of a ratio of about 0.003 with the 4422 or 3311 combinations. In addition to reducing how far the laser needs to be detuned from the transition, a larger $ \Omega_a/\Omega_m$ ratio also relaxes the intensity stability requirement, a key potential technical limitation. The position of the resonance depends on the laser power predominantly through the AC Stark shift on the molecular state, $ \Omega_m^2 / 2\Delta $. \textcolor{red}{KW: Can also include similar footnote here about cross coupling? The 2 in the denominator is assuming cross coupling and equal powers.} The ratio of the AC Stark shift to the Raman Rabi frequency is $ \Omega_m / \Omega_a $, Thus, the laser intensity needs to be stabilized to better than the inverse of this ratio to fluctuate by less than a linewidth in a coherent process. Thus, instead of a required intensity stability of $0.3 \% $ for the 4422 or 3311 combination, the 3322 combination only requires the intensity to be stabilized to $ 5\% $.  Thus, we choose the 3322 spin combination as our initial state and drive to the first bound state for the 3322 spin combination. (We might want to mention that those ratios are for 80 kHz spherical trap if we want to give more details about the coupled channel calculation anywhere...)

% In additional to the final and the excited state, it is also important to select the an initial atomic state in order to improve the coupling.
\begin{figure*}[ht!]
  \includegraphics[height=4.5cm]{fig2.pdf}
  \caption{
    (A) Geometry and polarization of trap and Raman beam relative to the bias magnetic field.
    The tweezer and Raman beam is focused through an objective to define the location of the
    atoms and molecule.
    We use a bias B field of $B_0=8.8 G$ alone the tweezer polarization
    to define the quantization axis.
    As a result, the atoms experiences predominately $\pi$ polarization from the tweezer.
    (B) Molecule formation pulse sequence. The tweezer initially consists of only up leg power.
    When driving the Raman transition, the up leg power is smoothly ramped down and
    the down leg power ramped up over $10\mu s$ while maintaining the total power of the tweezer.
    This minimizes the heating on the atoms due to power fluctuation while maximizes the time
    with maximum Raman Rabi frequency when the up and down leg powers are equal.
    (C) Raman resonance from atomic state $Na(2,2)\ Cs(3,3)$ to the first molecular bound state
    using a $0.1\mathrm{ms}$ pulse.
    The full width half maximum (FWHM) of $8.43(81) \mathrm{kHz}$ of the resonance
    is consistent with the FWHM for a coherent $0.1\mathrm{ms}$ $8.0 \mathrm{kHz}$ $\pi$ pulse.
    (D) Raman pulse time scan on resonance. A decaying Rabi oscillation can be observed
    proving the coherence of the Raman transfer process.
    \label{f-raman}}
\end{figure*}
